Bayesian exponential random graph modelling of KWL Burial network

### 2020-03-27 analysis/paper/paper.Rmd
paste some paragraphs from proposal and reorganize as a start

# Background

>>>>

Burials provide valuable information about social structures that reflect not only the personal life and status of the deceased, but also the expressions of their descendants, from which we are able to detect the social differences that can relate to social inequality and complexity (Ehrhardt 2009; Gamble and Zepeda 2002; Janes 2013; Johansen and Bauer 2015). Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow us to infer social structures, for examples, grave goods usually relate to social practices of trade, exchange, and gifting (Coward 2013: 252). One way to study the relationships between individual is network analysis, where burials represent actors, nodes in the network, linked by sharing the same attributes such as the same prestige goods.

Social Network Analysis (SNA) is a tool for understanding network structure and its effects on every part within the network (Freeman 2004; Salvini 2010). A social network contains a set of actors and a set of relationships connecting these actors, which could be groups, organizations or persons. Their relationships are flows of resources that reflect relations of control, dependence, and cooperation. In other words, actors are viewed as ‘nodes’ in a wider network, in which nodes are connected by multiple relationships, or ‘edges’, to other nodes (Coward 2013; Mizoguchi 2013; Potgieter, et al. 2008). SNA provides a way to visualize, quantify, and statistically test hypotheses about the relationships between actors. At Kiwulan, I will use individual burials as actors or nodes in the social network, and use SNA methods to test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions. Network ties will be drawn between burials with similar variables that can represent the same ‘community’ or social group. To measure change in social inequality, I will use the binary presence or absence of burial goods, the amount of prestige goods, grave dimensional attributes, and ordinal variables of age and sex categories to determine which actors have connections in the network (Sosna, et al. 2012).
Once the relationships between the actors have been described, I will compute statistical measures such as network degree centrality, graph density, and graph centralization. These measures are helpful to identify individual’s role and trends of network relationship, and I will use them to formally test hypotheses about the structure of the social network (Brughmans 2010: 280; Gjesfjeld 2014: 192). I predict that if there is a change from corporate to network mode following the European contact, then I expect the social network will show statistically significantly higher degrees of centrality and a higher density in the post-European period compared to pre-European period. This is because a social network graph that is highly centralized and very sparse indicates high concentrations of wealth, and higher social inequality.

Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities?
We ask whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an indigenous society. We use social network analysis (SNA) where burials represent actors (nodes in the network) linked by sharing the same prestige goods. Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities?

Exponential random graph models (ERGMs) are an important family of statistical models for analyzing network data and evaluating models of network structure. However, ERGMs are difficult to compute because their normalizing constant, which depends on model parameters, is intractable. A Bayesian framework allows for parameter inference using MCMC strategies which avoid the need for computationally intensive calculations of the normalizing constants. Using the R language, we use Bayesian SNA to study burials from Kiwulan, an Iron Age site in northeast Taiwan. We expect the network to present a higher degree of clustering (e.g. the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period. This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on indigenous groups. A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.

# Archaeological background

Based on the quantity and quality of grave goods among burials, we can reconstruct social networks from the distribution of prestige goods, including imported ceramic and beads based on the principle that prestige materials usually relate to social status.
